---
title: "Transcriptomics"
author: "Kamron Kimball"
date: "2/20/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Import today's data

Download airway_scaledcounts.csv and airway_metadata.csv to your R project.
Then:

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

We then want to take the average read counts of all the controlsand compare them to the average read counts of the treated cells.

First, we need to know how many genes we are dealing with.

```{r}
nrow(counts)
```

The we can calculte the averages counts per gene for the ontrol samples and for the treated samples. Afterwords, we compare those values to see if there is a difference overall between control and treated samples for a given gene.

This code allows us to access only the that are control within the metadata dataframe
```{r}
metadata$dex=="control"
control <- metadata[metadata$dex=="control",]
```

Now i can use the $id column of the control dataframe to access only the control columns in the counts dataframe, then sum and average all of those rows.
```{r}
control.mean <- rowSums(counts[,control$id])/nrow(control)
names(control.mean) <- counts$ensgene
```

Now we do the same for the treated values.

```{r}
metadata$dex=="treated"
treated <- metadata[metadata$dex=="treated",]
treated.mean <- rowSums(counts[,treated$id])/nrow(treated)
```

We shoud then combine these two dataframes

```{r}
meancounts <- data.frame(control.mean, treated.mean)
meancounts
```

We can make a scatterplot of this dataframe to visualize the values.

```{r}
plot(meancounts, type="p")
```

Making it a log plot will be more useful.

```{r}
plot(meancounts, type="p", log="xy")
```

We can calculate the log2foldchange of the dataframe to better express the data

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

We then have to modify this dataframe so that any of the genes that have a mean zero value are removed.

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

How many genes do we have left after modifying?

```{r}
nrow(mycounts)
```

We want to know how many points deviate from the normal line of what we calculated with the fold change.
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```
We can then sum to find how many deviations there are.
```{r}
sum(up.ind)
sum(down.ind)
```

##DESeq2 Analysis 
#Doing what we just did but the right way

```{r}
library(DESeq2)

```

Setup input for DESeq2 by specifying countdata, metadata and the design (i.e. which comparisons we are interested in)

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
res <- results(dds)
res
```


```{r}
plot(res$log2FoldChange, log(res$padj))
```

Making the plot visual more normal
 
```{r}
res$sig <- res$padj<0.05 & abs(res$log2FoldChange)>2
palette(c("gray","blue"))
plot(res$log2FoldChange, -log(res$padj), col=res$sig+1, ylab="-Log(P-value)", xlab="Log2(FoldChange)")
# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.1), col="darkgray", lty=2)

```

```{r}
# Setup our custom point color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

#Volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```







