---
title: "class09.Rmd"
author: "Kamron Kimball"
date: "2/4/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generate some example data for clustering
```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)

km <- kmeans(x,center=2,nstart=20)

```

Q. How many points are in each cluster?
Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?

```{r}
km$cluster
length(km$cluster)
table(km$cluster)
```

plot x colored by the kmeans cluster assignment and
    add cluster centers as blue points
```{r}
plot(x,col=km$cluster)
points(km$centers,col="blue")
```


## Hierarchical clustering in R

The main Hierarchical clustering function in R is caled 'hclust()'.
An important point is that you have to calculte the distance matrix deom your input data before calling 'hclust()'.

```{r}
# as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
# clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
```
hc
Call:
hclust(d = dist_matrix)
#Cluster method : complete
Distance : euclidean
#Number of objects: 60 

```{r}
#We will use our x again from above
d <- dist(x)
hc <- hclust(d)
plot(hc)
```

```{r}
# Our input is a distance matrix from the dist()
# function. Lets make sure we understand it first
dist_matrix <- dist(x)
dim(dist_matrix)
NULL
View( as.matrix(dist_matrix) )
dim(x)
#[1] 60 2
dim( as.matrix(dist_matrix) )
#[1] 60 60
# Note. symmetrical pairwise distance matrix
```

```{r}
# Create hierarchical cluster model: hc
hc <- hclust(dist(x))
# We can plot the results as a dendrogram
plot(hc)
# What do you notice?
# Does the dendrogram
# make sense based on
# your knowledge of x?
```

```{r}
# Draws a dendrogram
plot(hc)
abline(h=6, col="red", lty=2) 
abline(h=4, col="blue")
cutree(hc, h=6) #Cut by height h
```

To get cluster membership vector i need to "cut" the tree at a certain height to yield my separate cluster branches.

```{r}
cutree(hc, h=4)
```

```{r}
gp4 <-  cutree(hc, h=4)
table(gp4)
```

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters
Q. How does this compare to your known 'col' groups?

```{r}
z <- dist(x)
v <- hclust(z)

#Use "cutree" to get a cluster membership vector
ct <- cutree(v,h=2)

plot(v, col=ct)
table(ct)

#Make a plot that is grouped by ct
plot(x, col=ct)
```

```{r}
ukf <- read.csv("Data/UK_foods.csv", row.names=1)
ukf
barplot(as.matrix(ukf), beside=T, col=rainbow(nrow(ukf)))

```

a "Pairs" plot can be useful when we have small datasets like this one but often we are 

```{r}
pairs(ukf, col=rainbow(10), pch=16)
```

We can use PCA to better visualize this data

Principal Component analysis (PCA) with the "prcomp()" function
```{r}
# Use the prcomp() PCA function 
pca <- prcomp( t(ukf) )
summary(pca)
```

What is in my result object "pca"
```{r}
attributes(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(ukf), col=rainbow(4))
```







