---
title: "Class 12: Structural bioinformatics pt. 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics 

Here we inspect the types of structures in the main database

> Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. 

```{r}
#Read in file
stats <- read.csv("Data Export Summary.csv", row.names=1)


#calculate percentage values asked for
ans <- (stats$Total/sum(stats$Total))*100
names(ans) <- rownames(stats)
round(ans,2)
```

> Q2: Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub
document with working code that yields your answers.

```{r}
round((sum(stats$Proteins)/sum(stats$Total))*100, 2)
```


> Q3: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

1289


## Hands on Section 2


# Install packages in the R console not your Rmd
```{r}
# Install packages in the R console not your Rmd

install.packages("bio3d")
install.packages("ggplot2")
install.packages("ggrepel")
install.packages("devtools")
install.packages("BiocManager")

BiocManager::install("msa")
devtools::install_bitbucket("Grantlab/bio3d-view")
```


```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb
```

```{r}
attributes(pdb)
```

```{r}
pdb$atom
```

#Select protein only atoms

```{r}
prot <- atom.select(pdb, string = "protein", value = TRUE)
lig <- atom.select(pdb, string = "ligand", value = TRUE)


```

#And then write out these new PDB objects with the 'write.pdb()' function.

```{r}
write.pdb(lig, file="1hsg_ligand.pdb")
write.pdb(prot, file = "1hsg_protein.pdb")
```




