---
title: "Biological Network Analysis"
author: "Kamron Kimball"
date: "3/3/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Setup

We first have to setup the packages we need. These include **igraph** from CRAN and **RCy3** from bioconductor. Install these with **install.packages** and **BiocManager::install("RCy3")**.

We can then library these packages.

```{r}
library(RCy3)
library(igraph)
library(RColorBrewer)
```

We can also check our connection with cytoscape and test making a network and sending it to cytoscape.

```{r}
cytoscapePing()
g <- makeSimpleIgraph()
createNetworkFromIgraph(g,"myGraph")
```

We can include this Cytoscape rendered network image in our report with the following code:

```{r}
fig <- exportImage(filename="demo", type="png", height=350)
```

We can also customize the visuals of the network:
```{r}
setVisualStyle("Marquee")
```

These are all of the different styles available.

```{r}
styles <- getVisualStyleNames()
styles
```

The network can also be plotted within R itself:

```{r}
plot(g)
```

## Read our Data

We first have to read our data into R.

```{r}
## scripts for processing located in "inst/data-raw/"
prok_vir_cor <- read.delim("./virus_prok_cor_abundant.tsv", stringsAsFactors = FALSE)

## Have a peak at the first 6 rows
head(prok_vir_cor)
```

How many connections (i.e. edges) will we have in our network - lets look at the number of rows.

```{r}
nrow(prok_vir_cor)
```


The igraph package allows us to convert this dataframe into a network we can send to Cytoscape.

```{r}
g <- graph.data.frame(prok_vir_cor, directed = FALSE)
```

The plot of this data requires arguments to turn it into a readable network.

```{r}
plot(g)
plot(g, vertex.label=NA)
plot(g, vertex.size=3, vertex.label=NA)
```

When its readable, we can then send it to Cytoscape.

```{r}
createNetworkFromIgraph(g,"myIgraph")
```

## Network Community Detection

Lets cluster our network to find "community" structure...

Community structure detection algorithms try to find dense sub-graphs within larger network graphs (i.e. clusters of well connected nodes that are densely connected themselves but sparsely connected to other nodes outside the cluster) . Here we use the classic Girvan & Newman betweenness clustering method. The igraph package has lots of different community detection algorithms (i.e. different methods for finding communities).

```{r}
cb <- cluster_edge_betweenness(g)
cb
```

```{r}
plot(cb, y=g, vertex.label=NA,  vertex.size=3)
```

##Centrality analysis

Centrality gives an estimation on how important a node or edge is for the connectivity (or the information flow) of a network. It is a particularly useful parameter in signaling networks and it is often used when trying to find drug targets for example.

Centrality analysis often aims to answer the following question: Which nodes are the most important and why?

One centrality method that you can often find in publications is the Google PageRank score. For the explanation of the PageRank algorithm, see the following webpage: http://infolab.stanford.edu/~backrub/google.html

```{r}
pr <- page_rank(g)
head(pr$vector)
```

Lets plot our network with nodes size scaled via this page rank centrality scores.

```{r}
# Make a size vector btwn 2 and 20 for node plotting size

v.size <- BBmisc::normalize(pr$vector, range=c(2,20), method="range")
plot(g, vertex.size=v.size, vertex.label=NA)
```










